---
title: "cleanData"
author: "Sarah Wiegreffe"
date: "April 17, 2016"
output: html_document
---
Make sure your working directory is that of the project.
First, we import the datafile from the United Nation's High Commissioner for Refugees (exported from http://popstats.unhcr.org/en/persons_of_concern on March 23, 2016) and create a basic map.

```{r}
library(maps)

#import data from UNHCR
datAll=read.csv('./data/unhcr_popstats_export_persons_of_concern_all_data.csv', header = TRUE)

#basic map
wrd = map('world', fill=T)
```

We must reconcile the name differences between country names in the datafile vs. those in the map.
We only care about the destination countries (from the column "Country.of.asylum.or.residence" of our dataframe) matching the map, because the countries of origin will not be displayed on it. The resulting differences found by running this code were mapped my hand into the csv file "mapping_countries.csv".

```{r}
#remove ':'s from map names for the purpose of matching to dataset
mapChoices = unique(sapply(strsplit(wrd$names, ':', fixed=T), function(x) x[[1]]))

#get differences
setdiff(unique(datAll$Country.of.asylum.or.residence[]), mapChoices)

#look up missing values from the final mapChoices list with this command:
#subset(wrd$names, grepl(paste('China', collapse= "|"), wrd$names))

#map by hand in a csv file
```

This function cleans the dataset by converting columns to the correct datatype, and then converts all values in "Country.of.asylum.or.residence" that do not match the map country names to the correct name based off of the csv file. 

```{r}
#function definition
cleanData = function(df) {
  
  df_clean = df[, c("Year", "Country.of.asylum.or.residence", "Origin", "Refugees")]
  
  #Convert datatypes of columns away from factors
  df_clean$Country.of.asylum.or.residence = as.character(df$Country.of.asylum.or.residence)
  df_clean$Origin = as.character(df$Origin)
  df_clean$Refugees = as.numeric(df$Refugees)
  
  #fix mapping issues by changing some asylum country names in dataset to match 
  #those in map package's vocabulary.
  #import csv with mappings done by hand.
  nameUpdates = read.csv('./data/mapping_countries.csv', header = TRUE,
                         stringsAsFactors = FALSE)[,1:2]
  #str(nameUpdates)
  
  for (i in 1:dim(nameUpdates)[1]){
    df_clean$Country.of.asylum.or.residence[grepl(nameUpdates[i,1], 
                                                df_clean$Country.of.asylum.or.residence, fixed = TRUE)] = nameUpdates[i,2]
  }
  
  return(df_clean)
}
```

The last step is to call this function on the data we read in. Now we're ready to go on to the mapping and modeling!

```{r}
#call function
datCleaned = cleanData(datAll)

#Check that conversion process worked- should be none (old name)
datCleaned[datCleaned$Country.of.asylum.or.residence == 'United Rep. of Tanzania',]
#should be some rows now (new name)
datCleaned[datCleaned$Country.of.asylum.or.residence == 'Tanzania',]

#check that differences are reconciled- should return nothing except for any names containing ":"
setdiff(unique(datCleaned$Country.of.asylum.or.residence[]), mapChoices)

#check datatypes of columns
str(datCleaned)

#get unique origins, destinations, and years for inputs in app
origins = unique(datCleaned$Origin[])
destinations = unique(datCleaned$Country.of.asylum.or.residence[])
years = unique(datCleaned$Year[])

#start mapping
mapChoicesAll = sapply(strsplit(wrd$names, ':', fixed=T), function(x) x[[1]])
```